---
title: "Brainspan_vs_NullDistributions"
author: "Dylan Hughes"
date: "2023-08-23"
output: html_document
---

```{r}
knitr::asis_output(paste0("Last run:  ", date()))
```


```{r, include = F}
library(knitr)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)


library(dplyr)
library(ggplot2)
library(conflicted)
conflict_prefer('lmer', 'lmerTest')
conflict_prefer('recode', 'dplyr')
conflict_prefer('filter','dplyr')


library(devtools)
devtools::document('~/Documents/GitHub/dylanfuncs/dylanfuncs')
library(lmerTest)
library(gt)
library(corrplot)

pullModNum <- function(x, first) {
  for (i in c(2:length(x))) {
    split = strsplit(x[i], split = '')[[1]]
    num = as.numeric(split[(length(split)-1):length(split)])
    num.paste = paste(num[!is.na(num)], collapse = '')
    x[i] = paste('M', num.paste, collapse = '', sep = '')
  }
  
  final <- c(first, x[2:length(x)])
  return(final)
}
```

```{r, echo = F}
plotActualOnNull <- function(module, nulls, observed, snpcounts, numsnp = T) {
  pgsLab = substr(snpcounts$V1[1], 1, which(strsplit(snpcounts$V1[1], split = '')[[1]] == "_") - 1)
  
  if (grepl("EUR", pgsLab)) {
    pop = "EUR"
  } else if (grepl("AMR", pgsLab)) {
    pop = "AMR"
  } else if (grepl("AFR", pgsLab)) {
    pop = "AFR"
  } else {
    stop("Error with plotActualOnNull func: no population identifier in snpcount file")
  }
  
  if (grepl("02", pgsLab)) {
    find = paste0(pop, "02")
  } else {
    find = pop
  }
  
  pgs = sub(find, "", pgsLab)
  
  pgsLab = paste0(pgs, pop)
  
  if (numsnp == T) {
    title = paste0(pgsLab, ", ", module, ", # snps = ", snpcounts$V2[which(snpcounts$V3 == module)])
  } else {
    title = paste0(pgsLab, ", ", module)
  }
  
  plot <- ggplot(nulls[[module]], aes(x = PQB.Beta)) + 
    geom_histogram() + 
    geom_vline(xintercept = observed$z.pqb[which(rownames(observed) == module)], color = 'red') +
    ggtitle(title)
  return(plot)
}

loopModuleNulls <- function(nulls, observed, snpcounts, numsnp = T) {
  mods <- names(nulls)
  std = 8.740959
  
  pvals = rep(NA, length(mods))
  names(pvals) = mods
  
  for (m in mods) {
    nulls[[m]]$PQB.Beta = nulls[[m]]$PQB.Beta/std
    plot <- plotActualOnNull(module = m, nulls = nulls, observed = observed, snpcounts = snpcounts, numsnp = numsnp)
    
    observed.eff = observed$z.pqb[rownames(observed)==m]
    median.null = median(nulls[[m]]$PQB.Beta)
    #diff = abs(abs(median.null) - abs(observed.eff))
    
    # if (median.null < 0) {
    #   low.tail = length(which(nulls[[m]]$PQB.Beta < (median.null - diff)))
    # } else if (median.null > 0) {
    #   high.tail = length(which(nulls[[m]]$PQB.Beta > (median.null + diff)))
    # }
    if (observed.eff < 0) {
      tail = length(which(nulls[[m]]$PQB.Beta < observed.eff))
    } else if (observed.eff > 0) {
      tail = length(which(nulls[[m]]$PQB.Beta > observed.eff))
    }
    
    
   
    pval = tail/10000
    
    midpt.x = (median.null + max(nulls[[m]]$PQB.Beta)) / 2
  
    final.plot <- plot + 
      annotate("text", y = 700, x = midpt.x, label = paste0('1-sided p = ', pval)) + 
      xlab("STD Beta") + 
      ylab("Count")
    print(final.plot)
    
    pvals[which(mods==m)] = pval
  }
  
  return(pvals)
}

competitiveP <- function(nulls, observed) {
  
  mods = names(nulls)
  
  compP = rep(NA,length(mods))
  
  for (mod in mods) {
    observedP = observed$peas$z.pqb[which(rownames(observed$peas) == mod)]
    df = nulls[[mod]]
    compP[which(mods == mod)] = (length(which(df$PQB.P < observedP)) + 1) / (nrow(df) + 1)
  }
  names(compP) = mods
  return(compP)
}

readNulls <- function(pgs, pop) {
  nulls = list()
  for (mod in paste0('M', 1:18)) {
    file = paste0('~/Documents/ple_attention/brainspan_permutation_pipeline/', pgs,'_Results/vsBrainspanSNPs/', pop,'/phi02_refCorrected/', mod, '_', pgs, pop, '02Results_BrainspanSNPs.csv')
  
    if (file.exists(file)) {
      nulls[[mod]] = read.csv(file)
    }
  }
  
  return(nulls)
}

readsnpCounts <- function(pgs, pop) {
  
  if (paste0(pgs, pop) %in% c("ADHDEUR", "CPEUR")) {
    file = paste0('~/Documents/abcd/genetics/PGS/', pgs,'/',pop,'/',pgs,pop,'_Brainspan_SNPcounts.csv')  
  } else {
    file = paste0('~/Documents/abcd/genetics/PGS/', pgs,'/',pop,'/',pgs,pop,'02_Brainspan_SNPcounts.csv')
  }
  
  if (file.exists(file)) {
    out = read.csv(file)
  } else {
    stop(paste0("Could not find file: ", file))
  }
  
  return(out)
}

```

```{r, warning=F, echo = F, message=FALSE}
master <- read.csv('~/Documents/ple_attention/data/masterSheet_20240127.csv') %>%
  mutate(time.point = dplyr::recode(eventname, "baseline_year_1_arm_1" = 0, "1_year_follow_up_y_arm_1" = 1, "2_year_follow_up_y_arm_1" = 2, "3_year_follow_up_y_arm_1" = 3, "4_year_follow_up_y_arm_1" = 4))

master$rel_family_id = as.factor(master$rel_family_id)
master$site_id_l = as.factor(master$site_id_l)
master$z.pqb = as.numeric(scale(master$DistressScoreSum))

cognitive.vars = c('iiv_composite_zscore')
all.dvs = c(cognitive.vars, 'z.pqb')
short.dvs = substr(all.dvs, start = 1, stop = 15)

pgs = c('z.ADHDEUR.prscs.pgs','z.ADHDAMR02.prscs.pgs','z.ADHDAFR02.prscs.pgs',
        'z.CPEUR.prscs.pgs','z.CPAMR02.prscs.pgs','z.CPAFR02.prscs.pgs',
        'z.SCZEUR02.prscs.pgs','z.SCZAMR02.prscs.pgs','z.SCZAFR02.prscs.pgs', 
        'z.NDVEUR02.prscs.pgs','z.NDVAMR02.prscs.pgs','z.NDVAFR02.prscs.pgs')

covariates.simple = c("PC1","PC2","PC3","PC4","PC5",'interview_age_new','demo_sex_v2','(1|site_id_l/rel_family_id/subjectkey)')

covariates.full = c("PC1","PC2","PC3","PC4","PC5",'interview_age_new','demo_sex_v2','parentalEducation_new','income_recode','(1|site_id_l/rel_family_id/subjectkey)')

std = sd(master$DistressScoreSum, na.rm=T)
```


# Brainspan ADHD. 

### Reading in null distributions (10k permutations) and superimposing observed effects
Red vertical line indicates observed effect. Outcome of both permutations and observed effects is PQB.  
Value printed to the right of distribution represents the p-value of the permutation i.e., the proportion of values that are less than the observed effect (or greater than depending on the position of the observed effect relative to the median of the distribution). 


### EUR

```{r, include = F}
all.dvs <- all.dvs
adhdbrain.pgs <- colnames(select(master, z.ADHDEUR.prscs.pgs, contains('z.ADHDEUR_M')))
row.names=pullModNum(x = adhdbrain.pgs, first = "FG-ADHD")
col.names = substr(all.dvs, start = 1, stop = 15)

adhdbrainspan.eur <- loopModThroughVars(dvs = all.dvs, ivs = adhdbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

#write.csv(adhdbrainspan$betas, '~/Documents/abcd/genetics/PGS/brainspan_effects/ADHD_Brainspan_effects.csv', row.names=F)
```

### Table of effects on IIV and PQB where numbers represent beta coefficients, single stars represent uncorrected significance, and double stars FDR significance
```{r, echo = F}
adhdtable <- makeIntoTable(adhdbrainspan.eur)
adhdtable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
adhdeur_nulls = readNulls("ADHD", "EUR")

adhdeur_snpcounts <- readsnpCounts(pgs = "ADHD", pop = "EUR")
```


Modules for which observed effect is significantly larger than the null: M5
```{r, echo = F}
adhd_actual_effects = adhdbrainspan.eur$betas

adhdeur_NullP = loopModuleNulls(nulls = adhdeur_nulls, observed = adhd_actual_effects, snpcounts = adhdeur_snpcounts, numsnp = F)

adhdeur_CompP = competitiveP(adhdeur_nulls, adhdbrainspan.eur)
```

Test M5 for CIs
```{r}
model <- lmer(z.pqb ~ z.ADHDEUR_M5 + PC1 + PC2 + PC3 + PC4 + PC5 + interview_age_new + demo_sex_v2 + (1|site_id_l/rel_family_id/subjectkey), master)

gtLmerInterpret(model, var = "z.ADHDEUR_M5")
```


### AMR. 

M16 significantly negatively predicts PQB, but only be permutation. not by multiple comparison

```{r, include = F}
adhdbrain.pgs <- colnames(select(master, z.ADHDAMR02.prscs.pgs, contains('z.ADHDAMR02_M')))

adhdbrainspan.amr <- loopModThroughVars(dvs = all.dvs, ivs = adhdbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```

### Table of effects on IIV and PQB where numbers represent beta coefficients, single stars represent uncorrected significance, and double stars FDR significance
```{r, echo = F}
adhdtable <- makeIntoTable(adhdbrainspan.amr)
adhdtable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
adhdamr_nulls = readNulls("ADHD", "AMR")

adhdamr_snpcounts <- readsnpCounts(pgs = "ADHD", pop = "AMR")
```

```{r, echo = F}
adhd_actual_effects = adhdbrainspan.amr$betas

adhdamr_NullP = loopModuleNulls(nulls = adhdamr_nulls, observed = adhd_actual_effects, snpcounts = adhdamr_snpcounts)

adhdamr_CompP = competitiveP(nulls = adhdamr_nulls, observed = adhdbrainspan.amr)
```


### AFR. 

```{r, include = F}
adhdbrain.pgs <- colnames(select(master, z.ADHDAFR02.prscs.pgs, contains('z.ADHDAFR02_M')))

adhdbrainspan.afr <- loopModThroughVars(dvs = all.dvs, ivs = adhdbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```

### Table of effects on IIV and PQB where numbers represent beta coefficients, single stars represent uncorrected significance, and double stars FDR significance
```{r, echo = F}
adhdtable <- makeIntoTable(adhdbrainspan.afr)
adhdtable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
adhdafr_nulls = readNulls("ADHD", "AFR")

adhdafr_snpcounts <- readsnpCounts("ADHD", "AFR")
```


```{r, echo = F}
adhd_actual_effects = adhdbrainspan.afr$betas

adhdafr_NullP = loopModuleNulls(nulls = adhdafr_nulls, observed = adhd_actual_effects, snpcounts = adhdafr_snpcounts)

adhdafr_CompP = competitiveP(adhdafr_nulls, adhdbrainspan.afr)
```
<hr style="border:dotted">

# Brainspan CP. 


### EUR. 

10/23/2023: the output has 9.95k permutations. Will have to run the last 50 permutations before publishing

Table of effects on IIV and PQB where numbers represent beta coefficients, single stars represent uncorrected significance, and double stars FDR significance
```{r, include = F}
all.dvs <- all.dvs
cpbrain.pgs <- c('z.CPEUR.prscs.pgs',colnames(select(master, contains('z.CPEUR_M'))))
row.names=pullModNum(x = cpbrain.pgs, first = "FS-CP")
col.names = substr(all.dvs, start = 1, stop = 15)

cpbrainspan.eur <- loopModThroughVars(dvs = all.dvs, ivs = cpbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```

```{r, echo = F}
cptable <- makeIntoTable(cpbrainspan.eur)
cptable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
cpeur_nulls = readNulls("CP", "EUR")

cpeur_snpcounts <- readsnpCounts("CP", "EUR")
```

Red vertical line indicates observed effect. Outcome of both permutations and observed effects is PQB.  

Value printed to the right of distribution represents the p-value of the permutation i.e., the proportion of values that are less than the observed effect (or greater than depending on the position of the observed effect relative to the median of the distribution).  

Modules for which observed effect is significantly larger than the null: **M5** *observed effect significantly predicts PQB as well, but is not a candidate for mediation as there is no effect on attention. There may be other domains of cognition through which these genes influence PQB pathways*
```{r, echo = F}
cp_actual_effects = cpbrainspan.eur$betas

cpeur_NullP = loopModuleNulls(nulls = cpeur_nulls, observed = cp_actual_effects, snpcounts = cpeur_snpcounts)

cpeur_CompP = competitiveP(cpeur_nulls, cpbrainspan.eur)
```


### AMR. 

M15 significantly positively predicts PQB by permutation, not by initial multiple comparisons though

```{r, include = F}
cpbrain.pgs <- c('z.CPAMR02.prscs.pgs',colnames(select(master, contains('z.CPAMR02_M'))))

cpbrainspan.amr <- loopModThroughVars(dvs = all.dvs, ivs = cpbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```

```{r, echo = F}
cptable <- makeIntoTable(cpbrainspan.amr)
cptable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
cpamr_nulls = readNulls("CP", "AMR")

cpamr_snpcounts <- readsnpCounts("CP", "AMR")
```


```{r, echo = F}
cp_actual_effects = cpbrainspan.amr$betas

cpamr_NullP = loopModuleNulls(nulls = cpamr_nulls, observed = cp_actual_effects, snpcounts = cpamr_snpcounts)

cpamr_CompP = competitiveP(cpamr_nulls, cpbrainspan.amr)
```


### AFR.  

M3 and M10 significantly positively and negatively (respectively) predict PQB by permutation. not by initial multiple comparisons.

```{r, include = F}
cpbrain.pgs <- c('z.CPAFR02.prscs.pgs',colnames(select(master, contains('z.CPAFR02_M'))))

cpbrainspan.afr <- loopModThroughVars(dvs = all.dvs, ivs = cpbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```

```{r, echo = F}
cptable <- makeIntoTable(cpbrainspan.afr)
cptable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
cpafr_nulls = readNulls("CP", "AFR")

cpafr_snpcounts <- readsnpCounts("CP", "AFR")
```

```{r, echo = F}
cp_actual_effects = cpbrainspan.afr$betas

cpafr_NullP = loopModuleNulls(nulls = cpafr_nulls, observed = cp_actual_effects, snpcounts = cpafr_snpcounts)

cpafr_CompP = competitiveP(cpafr_nulls, cpbrainspan.afr)
```

<hr style="border:dotted">

# Brainspan SCZ


### EUR

Table of effects on IIV and PQB where numbers represent beta coefficients, single stars represent uncorrected significance, and double stars FDR significance. 

Modules for which observed effect is significantly larger than the null: **M6** *this module also significantly predicts PQB in models, but is not a candidate for mediation as it does not act on attention.*  
```{r, include = F}
all.dvs <- all.dvs
sczbrain.pgs <- c('z.SCZEUR02.prscs.pgs',colnames(select(master, contains('z.SCZEUR02_M'))))
row.names = pullModNum(x = sczbrain.pgs, first = "FG-SCZ")
col.names = substr(all.dvs, start = 1, stop = 15)

sczbrainspan.eur <- loopModThroughVars(dvs = all.dvs, ivs = sczbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```


```{r, echo = F}
scztable <- makeIntoTable(sczbrainspan.eur)
scztable$pretty.table
```


Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
sczeur_nulls = readNulls("SCZ", "EUR")

sczeur_snpcounts <- readsnpCounts("SCZ", "EUR")
```


Red vertical line indicates observed effect. Outcome of both permutations and observed effects is PQB.  

Value printed to the right of distribution represents the p-value of the permutation i.e., the proportion of values that are less than the observed effect (or greater than depending on the position of the observed effect relative to the median of the distribution). 

```{r, echo = F}
scz_actual_effects = sczbrainspan.eur$betas

sczeur_NullP = loopModuleNulls(nulls = sczeur_nulls, observed = scz_actual_effects, snpcounts = sczeur_snpcounts)

sczeur_CompP = competitiveP(sczeur_nulls, sczbrainspan.eur)
```


### AMR. 

```{r, include = F}
sczbrain.pgs <- c('z.SCZAMR02.prscs.pgs',colnames(select(master, contains('z.SCZAMR02_M'))))
row.names = pullModNum(x = sczbrain.pgs, first = "FG-SCZ")
col.names = substr(all.dvs, start = 1, stop = 15)

sczbrainspan.amr <- loopModThroughVars(dvs = all.dvs, ivs = sczbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```

```{r, echo = F}
scztable <- makeIntoTable(sczbrainspan.amr)
scztable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
sczamr_nulls = readNulls("SCZ", "AMR")

sczamr_snpcounts <- readsnpCounts("SCZ", "AMR")
```


```{r, echo = F}
scz_actual_effects = sczbrainspan.amr$betas

sczamr_NullP = loopModuleNulls(nulls = sczamr_nulls, observed = scz_actual_effects, snpcounts = sczamr_snpcounts)

sczamr_CompP = competitiveP(sczamr_nulls, sczbrainspan.amr)
```


### AFR. No significant modules, so no permutations

```{r, include = F}
sczbrain.pgs <- c('z.SCZAFR02.prscs.pgs',colnames(select(master, contains('z.SCZAFR02_M'))))

sczbrainspan.afr <- loopModThroughVars(dvs = all.dvs, ivs = sczbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)

```

```{r, echo = F}
scztable <- makeIntoTable(sczbrainspan.afr)
scztable$pretty.table
```

```{r, echo = F, eval = F}
scz_actual_effects = sczbrainspan.afr$betas

sczafr_NullP = ""

sczafr_CompP = ""
```


<hr style="border:dotted">

# Brainspan NDV. 


### EUR. 

10/24/23: still need to do 50 permutations before publishing

Table of effects on IIV and PQB where numbers represent beta coefficients, single stars represent uncorrected significance, and double stars FDR significance. 


```{r, include = F}
all.dvs <- all.dvs
ndvbrain.pgs <- c('z.NDVEUR02.prscs.pgs',colnames(select(master, contains('z.NDVEUR02_M'))))
row.names= pullModNum(x = ndvbrain.pgs, first = "FG-NDV")
col.names = substr(all.dvs, start = 1, stop = 15)

ndvbrainspan.eur <- loopModThroughVars(dvs = all.dvs, ivs = ndvbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)
```

```{r, echo = F}
ndvtable <- makeIntoTable(ndvbrainspan.eur)
ndvtable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
ndveur_nulls = readNulls("NDV", "EUR")

ndveur_snpcounts <- readsnpCounts("NDV", "EUR")
```

Red vertical line indicates observed effect. Outcome of both permutations and observed effects is PQB.  

Value printed to the right of distribution represents the p-value of the permutation i.e., the proportion of values that are less than the observed effect (or greater than depending on the position of the observed effect relative to the median of the distribution). 
```{r, echo = F}
ndv_actual_effects = ndvbrainspan.eur$betas

ndveur_NullP = loopModuleNulls(nulls = ndveur_nulls, observed = ndv_actual_effects, snpcounts = ndveur_snpcounts, numsnp = F)

ndveur_CompP = competitiveP(ndveur_nulls, ndvbrainspan.eur)
```

### AMR. 

```{r, include = F}
ndvbrain.pgs <- c('z.NDVAMR02.prscs.pgs',colnames(select(master, contains('z.NDVAMR02_M'))))

ndvbrainspan.amr <- loopModThroughVars(dvs = all.dvs, ivs = ndvbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)
```

```{r, echo = F}
ndvtable <- makeIntoTable(ndvbrainspan.amr)
ndvtable$pretty.table
```

Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
ndvamr_nulls = readNulls("NDV", "AMR")

ndvamr_snpcounts <- readsnpCounts("NDV", "AMR")
```


```{r, echo = F}
ndv_actual_effects = ndvbrainspan.amr$betas

ndvamr_NullP = loopModuleNulls(nulls = ndvamr_nulls, observed = ndv_actual_effects, snpcounts = ndvamr_snpcounts)

ndvamr_CompP = competitiveP(ndvamr_nulls, ndvbrainspan.amr)
```


### AFR. 

M6 significantly positively predicts PQB by permutation. not by initial multiple comparison.

```{r, include = F}
ndvbrain.pgs <- c('z.NDVAFR02.prscs.pgs',colnames(select(master, contains('z.NDVAFR02_M'))))

ndvbrainspan.afr <- loopModThroughVars(dvs = all.dvs, ivs = ndvbrain.pgs, rownames=row.names, colnames = col.names, covariates = covariates.simple, data = master, tstat = T)
```

```{r, echo = F}
ndvtable <- makeIntoTable(ndvbrainspan.afr)
ndvtable$pretty.table
```


Reading in null distributions (10k permutations) and superimposing observed effects. 
```{r}
ndvafr_nulls = readNulls("NDV", "AFR")

ndvafr_snpcounts <- readsnpCounts("NDV", "AFR")
```


```{r, echo = F}
ndv_actual_effects = ndvbrainspan.afr$betas

ndvafr_NullP = loopModuleNulls(nulls = ndvafr_nulls, observed = ndv_actual_effects, snpcounts = ndvafr_snpcounts)

ndvafr_CompP = competitiveP(ndvafr_nulls, ndvbrainspan.afr)

```

<hr style="border:dotted">


```{r, comparison of null effects, eval = F}
names(adhdeur_nulls)
names(adhdamr_nulls)
test <- t.test(adhdamr_nulls$M7$PQB.Beta, adhdeur_nulls$M7$PQB.Beta)

len = length(adhdamr_nulls) + length(adhdafr_nulls) + 
  length(cpamr_nulls) + length(cpafr_nulls) +
  length(sczamr_nulls) + 
  length(ndvamr_nulls) + length(ndvafr_nulls)

df <- data.frame("Module" = rep(NA, len),
                 "PGS" = rep(NA, len),
                 "Pop" = rep(NA, len),
                 "Mean_nonEur" = rep(NA, len),
                 "oneWayP" = rep(NA, len),
                 "Mean_Eur" = rep(NA, len),
                 "Dif" = rep(NA, len),
                 "Pval" = rep(NA, len))


count = 1
for (dis in c("adhd","cp","scz","ndv")) {
  for (pop in c("amr", "afr")) {
    
    if (dis == "scz" & pop == "afr") {
      next
    }
    
    nullName = paste0(dis, pop, "_nulls")
    modNames = names(eval(parse(text = nullName)))
    for (mod in modNames) {
      
      df$Module[count] = mod
      df$Pop[count] = pop
      df$PGS[count] = dis
      
      oneWay = eval(parse(text = paste0("t.test(", nullName, "$", mod, "$PQB.Beta)")))
      df$oneWayP[count] = oneWay$p.value
      
      eurName = paste0(dis, "eur_nulls")
      test <- eval(parse(text = paste0("t.test(", nullName, "$", mod, "$PQB.Beta, ", eurName, "$", mod, "$PQB.Beta)")))
      
      df$Mean_nonEur[count] = as.numeric(test$estimate[1])
      df$Mean_Eur[count] = as.numeric(test$estimate[2])
      df$Dif[count] = as.numeric(test$estimate[2]) - as.numeric(test$estimate[1])
      df$Pval[count] = test$p.value
      
      count = count + 1
    }
  }
}

df
```





```{r, compare module size with mean null effect in eur}
library(scales)

plotSnpCounts_onMeanNulls = function(population, len) {
  snpCounts = data.frame("Mod" = rep(NA, len),
                       "PGS" = rep(NA, len),
                       "Mean" = rep(NA, len),
                       "SE" = rep(NA, len),
                       "SNPCount" = rep(NA, len))

  count = 1
  for (dis in c("adhd","cp","scz","ndv")) {
    
    if (population == "afr" & dis == "scz") {
      next
    }
    
    moduleNames = names(eval(parse(text = paste0(dis, population, "_nulls"))))
    for (mod in moduleNames) {
    
      vector = eval(parse(text = paste0(dis, population, "_nulls$", mod, "$PQB.Beta")))
      snpCounts$Mod[count] = mod
      snpCounts$PGS[count] = toupper(dis)
      snpCounts$Mean[count] = mean(vector)
      snpCounts$SE[count] = sd(vector) / length(vector)
      
      snps = eval(parse(text = paste0(dis, "eur_snpcounts")))
      snpCounts$SNPCount[count] = snps[which(snps$V3 == mod), "V2"]
      
      count = count + 1
    }
  }
  
  plot <- ggplot(snpCounts, aes(x = SNPCount/1000, y = Mean, color = PGS)) + geom_point() + 
    geom_line(stat = 'smooth', alpha = 0.8) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    xlab("SNP Count (in thousands)") +
    ggtitle(toupper(population)) +
    geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE))
  
  print(plot)
  return(snpCounts)
}

len = length(adhdeur_nulls) + length(cpeur_nulls) + length(sczeur_nulls) + length(ndveur_nulls)
eurSnpCount = plotSnpCounts_onMeanNulls(population = "eur", len = len)

len = length(adhdamr_nulls) + length(cpamr_nulls) + length(sczamr_nulls) + length(ndvamr_nulls)
amrSnpCount = plotSnpCounts_onMeanNulls(population = "amr", len = len)

len = length(adhdafr_nulls) + length(cpafr_nulls) + length(ndvafr_nulls)
afrSnpCount = plotSnpCounts_onMeanNulls(population = "afr", len = len)
```

```{r}
adhd_nullvsobs <- eurSnpCount %>%
  filter(PGS == 'ADHD') 
adhd_nullvsobs$Mean = adhd_nullvsobs$Mean/std

adhdobs <- data.frame("Mod" = paste0("M", 1:18),
                      "PGS" = rep("ADHD", 18),
                      "Mean" = adhdbrainspan.eur$betas$z.pqb[2:nrow(adhdbrainspan.eur$betas)],
                      "SE" = rep(NA, 18),
                      "SNPCount" = adhd_nullvsobs$SNPCount)

adhd_nullvsobs2 = rbind(adhd_nullvsobs, adhdobs)
adhd_nullvsobs2$type = c(rep("Null", 18), rep("Obs", 18))

adhd_nullvsobs2$Mod = factor(adhd_nullvsobs2$Mod, levels = unique(arrange(adhd_nullvsobs, SNPCount)$Mod))

ggplot(adhd_nullvsobs2, aes(x = Mod, y = Mean, color = type)) + 
  geom_point() +
  geom_line(stat = 'smooth', alpha = 0.8) +
  theme(axis.text.x = element_text(angle = 45, face = 'bold'))
```


```{r}
for (i in names(adhdeur_nulls)) {

  if (i %in% names(ndveur_nulls)) {
    print(i)
    print(t.test(adhdeur_nulls[[i]]["PQB.Beta"], ndveur_nulls[[i]]["PQB.Beta"]))
  } 
  
}
```

<hr style="border:dotted">

# Competitive P results
```{r, comp-p-table}
compP.table <- as.data.frame(matrix("-",ncol = 12, nrow = 18))
colnames(compP.table) = c("ADHDEUR", "NDVEUR","SCZEUR","CPEUR",
                          "ADHDAMR", "NDVAMR","SCZAMR","CPAMR",
                          "ADHDAFR","NDVAFR","SCZAFR","CPAFR")
rownames(compP.table) = paste0("M", 1:18)

nullP.table <- as.data.frame(matrix("-", ncol = 12, nrow = 18))
colnames(nullP.table) = colnames(compP.table)
rownames(nullP.table) = rownames(compP.table)

for (pgs in colnames(compP.table)) {
  lower = tolower(pgs)
  
  pop = tolower(substr(pgs, start = (nchar(pgs)-2), stop = nchar(pgs)))
  pgsonly = tolower(sub(toupper(pop), "", pgs))
  
  rawP = eval(parse(text = paste0(pgsonly, "brainspan.", pop, "$peas$z.pqb[-1]")))
  rawP_Pass = which(rawP < 0.05)
  rawP_PassMods = paste0("M", rawP_Pass)
  
  fdrP = eval(parse(text = paste0(pgsonly, "brainspan.", pop, "$fdr$z.pqb[-1]")))
  fdrPass = which(fdrP < 0.05)
  fdrPassMods = paste0("M", fdrPass)

  
  CompPeas = eval(parse(text = paste0(lower, "_CompP")))
  NullPeas = eval(parse(text = paste0(lower, "_NullP")))
  
  compP.table[rawP_Pass, pgs] = round(as.numeric(CompPeas[which(names(CompPeas) %in% rawP_PassMods)]), 3)
  compP.table[fdrPass, pgs] = paste0(compP.table[fdrPass, pgs], "*")
  
  
  nullP.table[rawP_Pass, pgs] = round(as.numeric(NullPeas[which(names(CompPeas) %in% rawP_PassMods)]), 3)
  nullP.table[fdrPass, pgs] = paste0(nullP.table[fdrPass, pgs], "*")
  
  
  colnames(compP.table)[which(colnames(compP.table) == pgs)] = paste(toupper(pop), toupper(pgsonly), sep = "_")
  colnames(nullP.table)[which(colnames(nullP.table) == pgs)] = paste(toupper(pop), toupper(pgsonly), sep = "_")
  
}

gt(compP.table, rownames_to_stub = T) %>%
  tab_spanner_delim(
    delim = "_"
  ) %>%
  tab_style(
    style = cell_text(weight = 'bold', font = "Times New Roman"),
    locations = list(
      cells_stub(),
      cells_column_spanners())
   ) %>%
  tab_style(style = cell_text(font = "Times New Roman"),
            locations = list(cells_body(columns = everything(), rows = everything()), cells_stub())) %>%
    tab_style(style = cell_text(font = "Times New Roman"),
            locations = cells_column_labels()) %>%
  tab_footnote(
    footnote = "* = corresponding observed associations significant after FDR correction",
  ) %>%
  tab_style(style = cell_text(font = 'Times New Roman'),
            locations = cells_footnotes()) %>%
  tab_header(title = "Competitive P-values") %>%
  tab_style(style = cell_text(font = "Times New Roman"),
            locations = cells_title()) %>%
  gtsave(filename = "~/Documents/ple_attention/figures/competitive_P_table.png")
  


gt(nullP.table, rownames_to_stub = T) %>%
  tab_spanner_delim(
    delim = "_"
  ) %>%
  tab_style(
    style = cell_text(weight = 'bold', font = "Times New Roman"),
    locations = list(
      cells_stub(),
      cells_column_spanners())
   ) %>%
  tab_style(style = cell_text(font = "Times New Roman"),
            locations = list(cells_body(columns = everything(), rows = everything()), cells_stub())) %>%
    tab_style(style = cell_text(font = "Times New Roman"),
            locations = cells_column_labels()) %>%
  tab_footnote(
    footnote = "* = corresponding observed associations significant after FDR correction",
  ) %>%
  tab_style(style = cell_text(font = 'Times New Roman'),
            locations = cells_footnotes()
  ) %>%
  tab_header(title = "One-sided permuted P-values") %>%
  tab_style(style = cell_text(font = "Times New Roman"),
            locations = cells_title()) %>%
  gtsave(filename = "~/Documents/ple_attention/figures/oneside_permuted_p_table.html")
```



Not shown: lasso regression or PLS to select modules that explain most variation ?
```{r, eval=F}
library(glmnet)
library(pls)

plsModel <- plsr(DistressScoreSum ~ z.NDV_newadhd_M1 + z.NDV_newadhd_M2 + z.NDV_newadhd_M3 + z.NDV_newadhd_M4 + z.NDV_newadhd_M5 + z.NDV_newadhd_M6 + z.NDV_newadhd_M7 + z.NDV_newadhd_M8 + z.NDV_newadhd_M9 + z.NDV_newadhd_M10 + z.NDV_newadhd_M11 + z.NDV_newadhd_M12 + z.NDV_newadhd_M13 + z.NDV_newadhd_M14 + z.NDV_newadhd_M15 + z.NDV_newadhd_M16 + z.NDV_newadhd_M17 + z.NDV_newadhd_M18 + z.pc1 + z.pc2 + z.pc3 + z.pc4 + z.pc5, scale = F, validation= "CV", data = pgs.y3)

summary(plsModel)





pgs.y3 <- filter(master, eventname == '3_year_follow_up_y_arm_1') %>%
  filter(!is.na(DistressScoreSum)) %>%
  filter(!is.na(z.NDV_newadhd_M1))
y = pgs.y3$DistressScoreSum
x = data.matrix(pgs.y3[,cpbrain.pgs[2:length(cpbrain.pgs)]])

cv_model <- cv.glmnet(x = x, y = y, alpha = 1)
best_lambda <- cv_model$lambda.min

best_model <- glmnet(x, y, alpha = 1)
coef(best_model)


lasso.mod <- cv.glmnet(x = )
allNDVMod = lmer(DistressScoreSum ~ z.NDV_newadhd_M1 + z.NDV_newadhd_M2 + z.NDV_newadhd_M3 + z.NDV_newadhd_M4 + z.NDV_newadhd_M5 + z.NDV_newadhd_M6 + z.NDV_newadhd_M7 + z.NDV_newadhd_M8 + z.NDV_newadhd_M9 + z.NDV_newadhd_M10 + z.NDV_newadhd_M11 + z.NDV_newadhd_M12 + z.NDV_newadhd_M13 + z.NDV_newadhd_M14 + z.NDV_newadhd_M15 + z.NDV_newadhd_M16 + z.NDV_newadhd_M17 + z.NDV_newadhd_M18 + z.pc1 + z.pc2 + z.pc3 + z.pc4 + z.pc5 + interview_age_new + demo_sex_v2 + (1|site_id_l/rel_family_id/subjectkey), master)
summary(allNDVMod)
```

